---
title: "Appendix Figures for 'Estimating the effects of a California gun control program with Multitask Gaussian Processes'"
author: "Eli Ben-Michael, David Arbour, Avi Feller, Alexander Franks, and Steven Raphael"
date: "July 2022"

output: 
  html_document: default
  self_contained: no
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dev = "pdf")
library(haven)
library(tidyverse)
library(tidybayes)
library(cmdstanr)
library(posterior)
library(lmtest)
library(broom)
library(readxl)
library(ggrepel)
library(patchwork)
library(gsynth)
library(augsynth)
options(digits=3)
options(dev = "pdf")
# opts_chunk$set(dev = "pdf")

knitr::opts_chunk$set(dev = "pdf")
knitr::opts_chunk$set(fig.path = "../figs/appendix/")
```

```{r}
start_year <- 1997
end_year <- 2016

t_int <- 2007

guns <- read_dta("ucrpanel2.dta")

crimes <- read_dta("ucrthrough2018.dta") %>%
    mutate(across(c(violent_crime, homicide, rape_legacy, rape_revised, robbery,
           aggravated_assault, property_crime, burglary, larceny,
           motor_vehicle_theft),
           as.integer)) %>%
    rename(State = state_abbr, rape_rate = rape_legacy_rate, 
           murder_rate = homicide_rate, violent_rate = violent_crime_rate,
           assault_rate = aggravated_assault_rate,
           property_rate = property_crime_rate,
           mvt_rate = motor_vehicle_theft_rate, 
           Population = population) %>%
    filter(State != "DC") %>%
    # drop rape rate after 2015 since it switches to new definition
    mutate(rape_rate = ifelse(year > 2016, NA, rape_rate),
           treated = State == "CA", trt = treated * (year >= 2007))

if(start_year < 1995) {
  crimes <- crimes %>% filter(State != "MS")
}


guns %>%
    mutate(treated = State == "CA", trt = treated * (year >= 2007)) %>%
    mutate_at(vars(contains("rate")), list(log = log)) -> guns

# merge in covariates
cov_data <- read_excel("covariates_2005.xls") %>%
  mutate(logincome = log10(income)) %>%
  rename_with( ~ str_c(., "_2005"), -c(state,year)) %>%
  select(-year)


guns %>%
  inner_join(cov_data, by = c("State" = "state")) -> guns


gunrate_states <- guns %>% 
    filter(year >= start_year, year <= end_year,
           State != "DC", is.na(gunrate)) %>%
    distinct(State) %>% pull(State)
guns %>%
  inner_join(cov_data, by = c("State" = "state")) -> guns


gunrate_states <- guns %>% 
    filter(year >= start_year, year <= end_year,
           State != "DC", is.na(gunrate)) %>%
    distinct(State) %>% pull(State)

```


```{r, utility_functions}

get_posterior_gauss <- function(fit_mtgp) {
  draws <- as_draws_matrix(fit_mtgp$draws())
  means <- subset(draws, variable = "f")
  ppds <- subset(draws, variable = "f_samples")
  sigs <- subset(draws, variable = "sigman")

  return(list(mean_fn=means, ppd=ppds, sd=sigs))
}

get_posterior_pois <- function(fit_mtgp) {
  draws <- as_draws_matrix(fit_mtgp$draws())
  means <- subset(draws, variable = "f")
  ppds <- subset(draws, variable = "f_samples")
  return(list(mean_fn=means, ppd=ppds))
}



organize_post_samples <- function(post, ppd, unit, time, data) {
  
  time <- enquo(time)
  unit <- enquo(unit)
  times <- data %>% distinct(!!time) %>% pull(!!time)
  units <- data %>% distinct(!!unit) %>% pull(!!unit)
  
  # get time-unit-sample pairs
  time <- as_label(time)
  unit <- as_label(unit)
  
  times_units_samples <- expand.grid(1:nrow(post), times, units)

  names(times_units_samples) <- c("sample", time, unit)
  times_units_samples$mean_fn <- c(post)
  times_units_samples$ppd <- c(ppd)

  return(left_join(times_units_samples,
                   data[, c("year", "State", "homicide", "Population")],
                   by=c("year", "State")))
  
}

```

```{r}
res_gaus <- readRDS("../results/mtgp_fits_homoskedastic.RDS")
res_pois <- readRDS("../results/mtgp_fits_poisson.RDS")
names(res_gaus) <- 1:length(res_gaus) - 1
names(res_pois) <- 1:length(res_pois) - 1
```



## Figure A.1: CausalImpact Estimates

```{r causalimpact_est}
library(CausalImpact)

# CausalImpact Estimates
crimes %>% 
  filter(year >=1997) %>% 
  select(year, State, murder_rate) %>% 
  pivot_wider(id_cols=year, names_from=State, values_from=murder_rate) %>% 
  as.data.frame() -> causalimpact_df

ci_no_years = causalimpact_df[, -1]
ci_no_years[,c("CA", state_names[-5])] -> usable
CA_df = data.frame(CA=usable[,1])
impact_1 <- CausalImpact(CA_df, c(1,10), c(11,22), model.args = list(niter = 5000, nseasons = 5))
plot(impact_1)
```




## Figure C.2: Balance for Ranks 0-7 (Poisson)

```{r pre_process}
ranks <- 0:7

posts_pois <- lapply(res_pois, get_posterior_pois)

post_pred_ranks_pois <- lapply(ranks, function(r) organize_post_samples(posts_pois[[r+1]]$mean_fn, 
                                                                        posts_pois[[r+1]]$ppd,
                                                                        unit = State, time = year, 
                                                                        data = crimes %>% filter(year >= start_year)))

names(post_pred_ranks_pois) <- ranks

```


```{r ca_pois_balance_raw}
bind_rows(
  post_pred_ranks_pois, .id = "Rank"
  ) %>%
  group_by(year, State, Rank) %>%
  filter(year < 2007) %>%
  filter(State == "CA") %>%
  summarise(
        post_pred_mean = mean(ppd),
        post_pred_upper = quantile(ppd, .975),
        post_pred_lower = quantile(ppd, .025),
        post_pred_upper50 = quantile(ppd, 0.75),
        post_pred_lower50 = quantile(ppd, 0.25)
      ) %>%
  inner_join(crimes, by = c("State", "year")) %>%
  ggplot(aes(x = year, group = State)) +
  geom_ribbon(aes(ymax = post_pred_upper/ (Population / 1e5), ymin = post_pred_lower/ (Population / 1e5)),
              fill = "grey50", alpha = .5) +
  geom_ribbon(aes(ymax = post_pred_upper50/ (Population / 1e5), ymin = post_pred_lower50/ (Population / 1e5)),
              fill = "grey20", alpha = .5) +
  geom_line(aes(y = post_pred_mean / (Population / 1e5)), lwd = 1.25) + #color = rgb(228,26,28, maxColorValue = 255)) +
  geom_line(aes(y = murder_rate), lwd = 1, color = rgb(228,26,28, maxColorValue = 255)) +
  # geom_vline(xintercept = 2007, lty = 2) +
  facet_wrap(~ Rank, labeller = label_both, ncol=4) +
  xlab("") +
  ylab("Homicides per 100,000") +
  # ggtitle("Observed and posterior predictive distribution of homicide counts") +
  ggtitle("Observed and posterior predictive dist for California") + 
  theme_bw()
```


## Fig C.3: Coverage vs. Rank

### Coverage Plot (Poisson)

```{r, coverage_plot}

coverage_fn <- function(...){
  summarize(...,  mse_true = 
              sqrt(mean((mean_fn/scaled_population - homicide/scaled_population)^2)), 
                  mse_ppd = sqrt(mean((mean_fn/scaled_population - ppd/scaled_population)^2))) }

compute_coverage <- function(rank) {
  fit <- res_pois[[rank]]
  post <- get_posterior_pois(fit)$mean_fn
  y_pred <- get_posterior_pois(fit)$ppd
  post_data <- organize_post_samples(post,
                                     y_pred,
                                     unit = State,
                                     time = year,
                                     data = crimes %>% filter(year >= start_year))

  coverage_by_state = post_data %>% 
    filter(!(State=="CA" & year > t_int)) %>% 
    mutate(scaled_population =  Population/1e5) %>% 
    group_by(year, State) %>% 
    summarize(q975=quantile(ppd, 0.975),
              q75=quantile(ppd, 0.75),
              q25=quantile(ppd, 0.25),
              q025=quantile(ppd, 0.025),
              homicide=median(homicide)) %>% 
    ungroup() %>% 
    group_by(State) %>% 
    summarize(`95%`=1-mean(homicide > q975 | homicide < q025),
              `50%`=1-mean(homicide > q75 | homicide < q25)) 
  coverage_by_state %>% 
    add_row(State="AVG", 
            `95%`=mean(coverage_by_state$`95%`),
            `50%`=mean(coverage_by_state$`50%`)) %>% 
    mutate(rank=rank)
}

coverage_results <- map_dfr(names(res_gaus), 
        function(x) compute_coverage(x)) %>% 
        mutate(rank = as.factor(rank))

coverage_results %>% filter(State=="AVG") %>% 
  pivot_longer(2:3, names_to="Type", values_to="Coverage") %>%
  ggplot() + geom_line(aes(x=as.numeric(rank)-1, y=Coverage, col=Type), size=1.5) +   geom_hline(yintercept=0.5, linetype="dashed") + 
  geom_hline(data=tibble(M=c("95%", "50%"), val=c(0.95, 0.5)), aes(yintercept=val, col=M), linetype="dashed") + 
  ylim(c(0, 1)) + theme_bw(base_size=16) + xlab("Rank") + ggtitle("Poisson")

```

### Coverage Plot (Gaussian)

```{r, coverage_plot_gauss}

coverage_fn <- function(...){
  summarize(...,  mse_true = 
              sqrt(mean((mean_fn/scaled_population - homicide/scaled_population)^2)), 
                  mse_ppd = sqrt(mean((mean_fn/scaled_population - ppd/scaled_population)^2))) }

compute_coverage <- function(rank) {
  fit <- res_gaus[[rank]]
  post <- get_posterior_pois(fit)$mean_fn
  y_pred <- get_posterior_pois(fit)$ppd
  post_data <- organize_post_samples(post,
                                     y_pred,
                                     unit = State,
                                     time = year,
                                     data = crimes %>% filter(year >= start_year))

  coverage_by_state = post_data %>% 
    filter(!(State=="CA" & year > t_int)) %>% 
    mutate(scaled_population =  Population/1e5) %>% 
    group_by(year, State) %>% 
    summarize(q975=quantile(ppd, 0.975),
              q75=quantile(ppd, 0.75),
              q25=quantile(ppd, 0.25),
              q025=quantile(ppd, 0.025),
              homicide=median(homicide/scaled_population)) %>% 
    ungroup() %>% 
    group_by(State) %>% 
    summarize(`95%`=1-mean(homicide > q975 | homicide < q025),
              `50%`=1-mean(homicide > q75 | homicide < q25)) 
  coverage_by_state %>% 
    add_row(State="AVG", 
            `95%`=mean(coverage_by_state$`95%`),
            `50%`=mean(coverage_by_state$`50%`)) %>% 
    mutate(rank=rank)
}

coverage_results <- map_dfr(names(res_gaus), 
        function(x) compute_coverage(x)) %>% 
        mutate(rank = as.factor(rank))

coverage_results %>% filter(State=="AVG") %>% 
  pivot_longer(2:3, names_to="Type", values_to="Coverage") %>%
  ggplot() + geom_line(aes(x=as.numeric(rank)-1, y=Coverage, col=Type), size=1.5) +   geom_hline(yintercept=0.5, linetype="dashed") + 
  geom_hline(data=tibble(M=c("95%", "50%"), val=c(0.95, 0.5)), aes(yintercept=val, col=M), linetype="dashed") + 
  ylim(c(0, 1)) + theme_bw(base_size=16) + xlab("Rank") + ggtitle("Gaussian")

```


## Fig C.4: Global factor (rank 5)
```{r, global_factor}
fit_mtgp <- res_gaus[["5"]]
start_year <- 1997
end_year <- 2018

global_avg <- as_draws_matrix(fit_mtgp$draws()) %>% spread_draws(L_global[N1, N2], z_global[N1], global_offset) %>%
  group_by(.draw) %>% group_split() %>% map_dfr(.f = function(x) {
  L_global <- matrix(x$L_global, nrow=max(x$N1), ncol=max(x$N2), byrow=TRUE)
  z_global <- x %>% filter(N2==1) %>% pull(z_global)
  tibble(.draw = x$`.draw`[1], Year = start_year:end_year, f_global = as.numeric(L_global %*% z_global) + x$global_offset[1])
})

global_avg %>% group_by(Year) %>% 
  summarize(mean_f = mean(f_global), q025 = quantile(f_global, 0.025), q975 = quantile(f_global, 0.975)) %>%
  ggplot(aes(x=Year)) +  
  geom_ribbon(aes(ymax = q975, ymin = q025),
              fill = "grey50", alpha = .5) +
  geom_line(aes(y=mean_f)) + theme_bw(base_size=16) +
  ylab("Homicides per 100,000") +
  ggtitle("Homicides per 100,000 (State Average)") +
  xlim(c(1997, 2016))
```


## Fig C.5: Individual factors

```{r, factor_analysis, cache = TRUE}
fit_mtgp <- res_gaus[[6]]

offsets <- as_draws_matrix(fit_mtgp$draws()) %>% spread_draws(state_offset[D])

Ltibble <- as_draws_matrix(fit_mtgp$draws()) %>% spread_draws(L_f[N1, N2])
Ztibble <- as_draws_matrix(fit_mtgp$draws()) %>% spread_draws(z_f[N1, K])
Wtibble <- as_draws_matrix(fit_mtgp$draws()) %>% spread_draws(k_f[K, D])
sigman <- as_draws_matrix(fit_mtgp$draws()) %>% spread_draws(sigman)
population <- crimes %>% group_by(State) %>% summarize(pop = mean(Population)/1e5) %>% pull(pop)

N <- max(Ltibble$N1)
K <- max(Ztibble$K)
D <- max(Wtibble$D)
ndraws <- max(Ltibble$`.draw`)
f_array <- array(0, dim=c(ndraws, N, K))
w_array <- array(0, dim=c(ndraws, K, D))

L <- matrix(Ltibble %>% filter(.draw == 2000) %>% pull(L_f), nrow=N, ncol=N, byrow=TRUE)
Z <- matrix(Ztibble %>% filter(.draw == 2000) %>% pull(z_f), nrow=N, ncol=K, byrow=TRUE)
W <- matrix(Wtibble %>% filter(.draw == 2000) %>% pull(k_f), nrow=K, ncol=D, byrow=TRUE)

Fbaseline <- svd(L %*% Z)$u

for(d in 1:ndraws) {
  # if(d %% 50 == 0)
  #   print(d)
  L <- matrix(Ltibble %>% filter(.draw == d) %>% pull(L_f), nrow=N, ncol=N, byrow=TRUE)
  Z <- matrix(Ztibble %>% filter(.draw == d) %>% pull(z_f), nrow=N, ncol=K, byrow=TRUE)
  W <- matrix(Wtibble %>% filter(.draw == d) %>% pull(k_f), nrow=K, ncol=D, byrow=TRUE)
  
  ones_vec <- cbind(rep(1/sqrt(N), N))
  svdF <- svd((diag(N) - ones_vec %*% t(ones_vec)) %*% L %*% Z)
  F <- svdF$u
  
  svd_res <- svd(t(Fbaseline) %*% F)
  R <-  svd_res$u %*% t(svd_res$v)
  f_array[d, ,] <- F %*% t(R)
  w_array[d, , ] <- R %*% diag(svdF$d) %*% t(svdF$v) %*% W
  
}


mean_factors <- apply(f_array, 2:3, mean)
colnames(mean_factors) <- paste("Factor", 1:dim(f_array)[3])
factor_tibble <- as_tibble(mean_factors)
factor_tibble$year <- start_year:end_year

factor_tibble %>% pivot_longer(cols=-year, names_to="Factor") %>% 
  ggplot() + geom_line(aes(x=year, y=value, col=Factor)) + theme_bw()

Wmean <- apply(w_array, 2:3, mean)
wsvd <- svd(Wmean)
Wmean <- diag(wsvd$d) %*% t(wsvd$v)
colnames(Wmean) <- unique(crimes$State)

mean_factors <- apply(f_array, 2:3, mean)
colnames(mean_factors) <- paste("Factor", 1:dim(f_array)[3])
mean_factors %*% wsvd$u

state_indices <- unique(crimes$State)

sgn <- sign(Wmean[, "CA"])
Wmean <- Wmean * sgn
w_sds <- apply(Wmean, 1, sd)
Wmean <- Wmean/w_sds
mean_factors <- t(sgn*t(mean_factors)*w_sds)
 
factor_tibble <- as_tibble(mean_factors)
factor_tibble$year <- start_year:end_year

factors_plot <- factor_tibble %>% pivot_longer(cols=-year, names_to="Factor") %>% 
  ggplot() + geom_line(aes(x=year, y=value, col=Factor)) +
  theme_bw() +
  geom_hline(yintercept=0, linetype="dashed") +
  xlim(c(1997, 2016))



ca_loadings_plot <- tibble(Loadings=Wmean[, "CA"], fact=colnames(mean_factors)) %>% ggplot(aes(x=fact, y=Loadings)) + geom_bar(stat="identity") + ylab("Loadings for California") + xlab("") + theme_bw() + coord_flip() 
ca_loadings_plot

factors_plot / ca_loadings_plot

```



## Fig C.6: Time weights
```{r}
fits = readRDS('../results/mtgp_fits_homoskedastic.RDS')


draws_mat = as_draws_matrix(fits[[6]]$draws()) 
draws_mat %>% spread_draws(sigman) -> sigman_df
draws_mat %>% spread_draws(lengthscale_f) -> ls_f_df
# factors
draws_mat %>% spread_draws(k_f[k_f, D]) -> k_f_df
# kernel matrix
draws_mat %>% spread_draws(K_f[A, B]) -> K_f


get_relevant_vars = function(iteration) {
  k_f = k_f_df %>% filter(.iteration==iteration) %>% pull('k_f') %>% matrix(5, 50)
  ones_vec = rep(1, 5)
  k_f = (diag(5) - ones_vec %*% t(ones_vec)) %*% k_f
  list(
    K = K_f %>% filter(.iteration==iteration) %>% pull('K_f') %>% matrix(22, 22),
    Theta=t(k_f) %*% k_f,
    sigma=sigman_df %>% filter(.iteration==iteration) %>% pull(sigman)
  )
}

get_weights = function(iteration) {
  if(iteration %% 5 == 0){
    print(iteration)
  }
  vars = get_relevant_vars(iteration)
  # get California in 2008
  solution = solve(
    kronecker(vars$K, vars$Theta) + sqrt(vars$sigma) * diag(1/pops), 
    kronecker(vars$K[11,], vars$Theta[5,])
  )
  sol_mat = matrix(solution, 50, 22)
  unit_weights = rowSums(sol_mat)
  
  sol = solve( kronecker(vars$K, vars$Theta) + sqrt(vars$sigma) * diag(1/pops), kronecker(vars$K[,11], vars$Theta))
  time_weights = colSums(matrix(rowSums(sol), 50, 22))
  
  return(list(time=time_weights, unit=unit_weights, mat=sol_mat))
}

weight_distributions = mclapply(1:4000, compute_weights, mc.cores=5)
# Time Weights
time_weights = sapply(1:4000, function(i) colSums(weight_distributions[[i]]))
values = data.frame(t(apply(time_weights, 1, function(d) quantile(d, c(0.05, 0.5, 0.95)))))
plot = ggplot(values, aes(x=times, y=fifty)) + 
  geom_line(lwd=0.9) + 
  geom_ribbon(aes(ymin=five, ymax=ninetyfive, x=times, fill = "band"), alpha = 0.3) + 
  theme_bw() + 
  scale_fill_manual("",values="grey12") + 
  ylab("Weight") + 
  xlab("Year") +
  theme(legend.position = "none") + 
  geom_vline(xintercept=2007, linetype="dashed", alpha=0.5)
pdf("time_weights.pdf")
print(plot)
dev.off()

```


## Fig C.7: Unit weights
```{r unit_weights}
# Plot the unit weights
unit_weights = sapply(1:4000, function(i) rowSums(weight_distributions[[i]]))

weight_df = data.frame(
  State=unique(crimes$State),
  t(apply(unit_weights, 1, function(col) quantile(col, c(0.05, 0.38, 0.5, .68, 0.95))))
)
names(weight_df) <- c("State", "five", "thirtyeight", "fifty", "sixtyeight", "ninetyfive")
plot = ggplot(weight_df, aes(x=reorder(State, fifty), y=fifty)) +
  geom_boxplot(fill="blue") +
  theme_bw() +
  theme(text = element_text(size=10), axis.line.x = "none") +
  geom_hline(yintercept=0, linetype='dashed', alpha=0.5) +
  geom_crossbar(aes(ymin=five, ymax=ninetyfive), width=0.1, linetype=0, fill="black", alpha=0.5) +
  geom_crossbar(aes(ymin=thirtyeight, ymax=sixtyeight), width=0.2, linetype=0, fill="black", alpha=0.6) +
  ylab("Unit Weight") + xlab("State") + coord_flip()
```


## Fig C.8: Impact estimates for Gaussian all ranks

```{r pre_process_gauss}
ranks <- 0:7

posts_gaus <- lapply(res_gaus, get_posterior_gauss)

post_pred_ranks_gaus <- lapply(ranks, function(r) organize_post_samples(posts_pois[[r+1]]$mean_fn, 
                                                                        posts_pois[[r+1]]$ppd,
                                                                        unit = State, time = year, 
                                                                        data = crimes %>% filter(year >= start_year)))

names(post_pred_ranks_gaus) <- ranks

```


```{r ca_plot_h_ranks}

bind_rows(
  post_pred_ranks_gaus, .id = "Rank"
  ) %>% 
  group_by(year, State, Rank)  %>%
  summarise(
        post_pred_mean = mean(ppd/Population*1e5),
        post_pred_upper = quantile(ppd/Population*1e5, .975),
        post_pred_lower = quantile(ppd/Population*1e5, .025)
      ) %>%
  inner_join(crimes, by = c("State", "year")) %>%
  filter(State == "CA") %>%
  ggplot(aes(x = year, group = State)) +
  geom_ribbon(aes(ymax = post_pred_upper, ymin = post_pred_lower),
              fill = "grey50", alpha = .5) +
  geom_line(aes(y = post_pred_mean), color = rgb(228,26,28, maxColorValue = 255)) +
  geom_line(aes(y = murder_rate)) +
  geom_vline(xintercept = 2007, lty = 2) +
  xlim(c(1997, 2016)) +
  facet_wrap(~ Rank, labeller = label_both, ncol=4) +
  xlab("") +
  ylab("Homicides per 100,000") +
  ggtitle("Observed and posterior predictive distribution of homicide rates") +
  theme_bw()
```




## Fig C.9: Impact estimates for Gaussian all ranks (covariate adjusted)

```{r pre_process_gaus_adj}
res_gaus_adj <- readRDS("../results/mtgp_fits_homoskedastic_linear_adj.RDS")

ranks <- 0:7

posts_gaus <- lapply(res_gaus_adj, get_posterior_gauss)

post_pred_ranks_gaus <- lapply(ranks, function(r) organize_post_samples(posts_pois[[r+1]]$mean_fn, 
                                                                        posts_pois[[r+1]]$ppd,
                                                                        unit = State, time = year, 
                                                                        data = crimes %>% filter(year >= start_year)))

names(post_pred_ranks_gaus) <- ranks

```


```{r ca_plot_gauss_ranks_adj}

bind_rows(
  post_pred_ranks_gaus, .id = "Rank"
  ) %>% 
  group_by(year, State, Rank)  %>%
  summarise(
        post_pred_mean = mean(ppd/Population*1e5),
        post_pred_upper = quantile(ppd/Population*1e5, .975),
        post_pred_lower = quantile(ppd/Population*1e5, .025)
      ) %>%
  inner_join(crimes, by = c("State", "year")) %>%
  filter(State == "CA") %>%
  ggplot(aes(x = year, group = State)) +
  geom_ribbon(aes(ymax = post_pred_upper, ymin = post_pred_lower),
              fill = "grey50", alpha = .5) +
  geom_line(aes(y = post_pred_mean), color = rgb(228,26,28, maxColorValue = 255)) +
  geom_line(aes(y = murder_rate)) +
  geom_vline(xintercept = 2007, lty = 2) +
  facet_wrap(~ Rank, labeller = label_both, ncol=4) +
  xlim(c(1997, 2016)) +
  xlab("") +
  ylab("Homicides per 100,000") +
  ggtitle("Observed and posterior predictive distribution of homicide rates") +
  theme_bw()
```









## Fig C.10: Impact estimates for Poisson all ranks

```{r ca_plot_pois_ranks}

bind_rows(
  post_pred_ranks_pois, .id = "Rank"
  ) %>%
  group_by(year, State, Rank) %>%
  summarise(
        post_pred_mean = mean(ppd),
        post_pred_upper = quantile(ppd, .975),
        post_pred_lower = quantile(ppd, .025)
      ) %>%
  inner_join(crimes, by = c("State", "year")) %>%
  filter(State == "CA") %>%
  ggplot(aes(x = year, group = State)) +
  geom_ribbon(aes(ymax = post_pred_upper/ (Population / 1e5), ymin = post_pred_lower/ (Population / 1e5)),
              fill = "grey50", alpha = .5) +
  geom_line(aes(y = post_pred_mean / (Population / 1e5)), color = rgb(228,26,28, maxColorValue = 255)) +
  geom_line(aes(y = murder_rate)) +
  geom_vline(xintercept = 2007, lty = 2) +
  facet_wrap(~ Rank, labeller = label_both, ncol=4) +
  xlab("") +
  ylab("Homicides per 100,000") +
  ggtitle("Observed and posterior predictive distribution of homicide counts") +
  theme_bw()
```











## Fig C.11: Impact estimates for Poisson all ranks (covariate adjusted)

```{r pre_process_adj}
res_pois_adj <- readRDS("../results/mtgp_fits_poisson_linear_adj.RDS")

ranks <- 0:7

posts_pois <- lapply(res_pois_adj, get_posterior_pois)

post_pred_ranks_pois <- lapply(ranks, function(r) organize_post_samples(posts_pois[[r+1]]$mean_fn, 
                                                                        posts_pois[[r+1]]$ppd,
                                                                        unit = State, time = year, 
                                                                        data = crimes %>% filter(year >= start_year)))

names(post_pred_ranks_pois) <- ranks

```

```{r ca_plot_pois_ranks_adj}

bind_rows(
  post_pred_ranks_pois, .id = "Rank"
  ) %>%
  group_by(year, State, Rank) %>%
  summarise(
        post_pred_mean = mean(ppd),
        post_pred_upper = quantile(ppd, .975),
        post_pred_lower = quantile(ppd, .025)
      ) %>%
  inner_join(crimes, by = c("State", "year")) %>%
  filter(State == "CA") %>%
  ggplot(aes(x = year, group = State)) +
  geom_ribbon(aes(ymax = post_pred_upper/ (Population / 1e5), ymin = post_pred_lower/ (Population / 1e5)),
              fill = "grey50", alpha = .5) +
  geom_line(aes(y = post_pred_mean / (Population / 1e5)), color = rgb(228,26,28, maxColorValue = 255)) +
  geom_line(aes(y = murder_rate)) +
  geom_vline(xintercept = 2007, lty = 2) +
  facet_wrap(~ Rank, labeller = label_both, ncol=4) +
  xlim(c(1997, 2016)) +
  xlab("") +
  ylab("Homicides per 100,000") +
  ggtitle("Observed and posterior predictive distribution of homicide counts") +
  theme_bw()
```



## Fig C.12: Comparison with other methods
```{r fit_alt_models}

crimes %>%
    filter(year >=start_year, year <= end_year, 
            State != "DC") %>%
    gsynth(murder_rate ~ trt, ., 
           index = c("State", "year"), se=FALSE, r = 2, CV = TRUE) -> gsyn


crimes %>%
    filter(year >=start_year, year <= end_year, 
            State != "DC") %>%
    gsynth(murder_rate ~ trt, ., index = c("State", "year"), 
           se=FALSE, estimator = "mc") -> mcp


crimes %>% 
    filter(year >= start_year, year <= end_year,
           State != "DC") %>%
    augsynth(murder_rate ~ treated,
                      State, year, t_int = t_int,
                      data = ., 
                      "None", scm = TRUE, fixedeff = TRUE) -> murder_descm

crimes %>% 
    filter(year >= start_year, year <= end_year,
           State != "DC") %>%
    augsynth(murder_rate ~ treated,
                      State, year, t_int = t_int,
                      data = ., 
                      "None", scm = TRUE, fixedeff = FALSE) -> murder_scm

crimes %>% 
    filter(year >= start_year, year <= end_year,
           State != "DC") %>%
    augsynth(murder_rate ~ treated,
                      State, year, t_int = t_int,
                      data = ., 
                      progfunc = "ridge", scm = TRUE, fixedeff = FALSE,
                      min_1se = FALSE) -> murder_ascm

```


```{r synth_estimates_murder_gsynth_gap, fig.height=6, fig.width=7,}

post_pred_ranks_pois$`5` %>% 
  group_by(year, State) %>%
  filter(State == "CA", year <= end_year) %>%
  # filter(Rank == 5) %>% 
  summarise(
        post_pred_mean = mean(ppd),
        post_pred_upper = quantile(ppd, .975),
        post_pred_lower = quantile(ppd, .025),
        post_pred_upper50 = quantile(ppd, 0.75),
        post_pred_lower50 = quantile(ppd, 0.25)
      ) %>% 
  inner_join(crimes, by = c("State", "year"))  %>%
  mutate(post_pred_mean = post_pred_mean / (Population / 1e5),
         post_pred_upper = post_pred_upper / (Population / 1e5),
         post_pred_lower = post_pred_lower / (Population / 1e5),
         post_pred_upper50 = post_pred_upper50 / (Population / 1e5),
         post_pred_lower50 = post_pred_lower50 / (Population / 1e5)) %>%
   mutate(post_pred_mean = murder_rate - post_pred_mean,
                     post_pred_upper = murder_rate - post_pred_upper,
                     post_pred_lower = murder_rate - post_pred_lower,
                     post_pred_upper50 = murder_rate - post_pred_upper50,
                     post_pred_lower50 = murder_rate - post_pred_lower50)  -> pois_rank5_posterior



bind_rows(data.frame(Estimate = gsyn$att) %>% 
            mutate(Time = start_year:end_year, Method = "GSynth"),
            data.frame(Estimate = mcp$att) %>% 
            mutate(Time = start_year:end_year, Method = "Matrix Completion"),
          summary(murder_ascm, inf = FALSE)$att %>% mutate(Method = "Ridge ASCM"),
          summary(murder_scm, inf = FALSE)$att %>% mutate(Method = "SCM"),
          summary(murder_descm, inf = FALSE)$att %>% mutate(Method = "Intercept-Shifted SCM")
          ) %>%
  rename(att=Estimate, year=Time) %>%
  mutate(label = case_when(
            Method == "GSynth" & year == end_year ~ Method,
            Method == "Intercept-Shifted SCM" & year == end_year - 1 ~ Method,
            Method == "Ridge ASCM" & year == end_year - 2 ~ Method,
            Method == "Matrix Completion" & year == end_year - 3 ~ Method,
            Method == "SCM" & year == end_year - 4 ~ Method
  )) %>%
  ggplot(aes(x = year)) +
  geom_ribbon(aes(ymax = post_pred_upper , ymin = post_pred_lower),
            fill = "grey50", alpha = .5, data = pois_rank5_posterior) +
  geom_ribbon(aes(ymax = post_pred_upper50 , ymin = post_pred_lower50 ),
            fill = "grey20", alpha = .5, data = pois_rank5_posterior) +
  geom_line(aes(y = post_pred_mean), lwd = 1.25, data = pois_rank5_posterior) + 
  geom_line(aes(y = att, color = Method, lty = Method), size=1) + 
  geom_vline(xintercept=2006.5, lty=2, size = 1) + 
  geom_hline(yintercept=0) + 
  geom_label_repel(aes(y = att, label = label, color = Method), nudge_x = -7) +
  # facet_wrap(~Method, ncol = 1) +
  xlab("Year") +
  ylab("Difference in Homicides per 100,000") +
  scale_color_brewer("", type = "qual", palette = "Set1") +
  scale_fill_brewer("", type = "qual", palette = "Set1") + 
  guides(color="none", fill="none", lty = "none") +
  theme_bw() +
  theme(panel.grid.minor.x=element_blank(),
        legend.position="bottom")
  
```




## Fig. C.13: Leave One Out Model Estimates

```{r state_loo_draws}
pois_loo <- readRDS("../results/pois_loo_draws.RDS")

state_order <- readRDS("../results/state_order.RDS")

lapply(pois_loo,
  function(x) {
    draws <- as_draws_matrix(x)
    means <- subset(draws, variable = "f")
    ppds <- subset(draws, variable = "f_samples")
    list(mean_fn=means, ppd=ppds)
  }) -> pois_loo


organize_post_samples <- function(post, ppd, unit, time, data) {
  
  time <- enquo(time)
  unit <- enquo(unit)
  times <- data %>% distinct(!!time) %>% pull(!!time)
  units <- data %>% distinct(!!unit) %>% pull(!!unit)
  
  # get time-unit-sample pairs
  time <- as_label(time)
  unit <- as_label(unit)
  
  times_units_samples <- expand.grid(1:nrow(post), times, units)

  names(times_units_samples) <- c("sample", time, unit)
  times_units_samples$mean_fn <- c(post)
  times_units_samples$ppd <- c(ppd)

  return(left_join(times_units_samples,
                   data[, c("year", "State", "homicide", "Population")],
                   by=c("year", "State")))
  
}

post_pred_pois_loo <- lapply(1:length(state_order),
  function(k) organize_post_samples(pois_loo[[k]]$mean_fn, 
                        pois_loo[[k]]$ppd,
                        unit = State, time = year, 
                        data = crimes %>% filter(State != !!state_order[k], year >= start_year)))

names(post_pred_pois_loo) <- state_order

end_year <- 2016

map_df(1:length(state_order),
  function(k) {
    post_pred_pois_loo[[k]]  %>%
      filter(State == "CA", year <= end_year, year >= 2007) %>%
      mutate(gap = (homicide - ppd) / Population * 1e5 ) %>%
      group_by(sample) %>%
      summarise(gap = mean(gap)) %>%
      ungroup() %>%
      summarise(post_mean = mean(gap),
                pq25 = quantile(gap, 0.25),
                pq2.5 = quantile(gap, 0.025),
                pq75 = quantile(gap, 0.75),
                pq97.5 = quantile(gap, 0.975)) %>%
      mutate(State = state_order[k])
  }) %>%
  mutate(loo = TRUE) %>%
  bind_rows(
    post_pred_ranks_pois$`5` %>%
      filter(State == "CA", year <= end_year, year >= 2007) %>%
      mutate(gap = (homicide - ppd) / Population * 1e5) %>%
      group_by(sample) %>%
      summarise(gap = mean(gap)) %>%
      ungroup() %>%
      summarise(post_mean = mean(gap),
                pq25 = quantile(gap, 0.25),
                pq2.5 = quantile(gap, 0.025),
                pq75 = quantile(gap, 0.75),
                pq97.5 = quantile(gap, 0.975)) %>%
      mutate(State = "", loo = FALSE)
  ) %>%
  mutate(State = factor(State, levels = c(rev(state_order), ""))) %>%
  ggplot(aes(x = State)) +
    geom_hline(yintercept = 0, lty = 2) +
  geom_segment(aes(y = pq2.5, yend = pq97.5, xend = State), lineend='round',
               size = .4, color = "grey50", data = . %>% filter(State != "")) +
  geom_segment(aes(y = pq25, yend = pq75, xend = State), lineend='round',
                data = . %>% filter(State != "")) +
  geom_point(aes(y = post_mean),  data = . %>% filter(State != "")) +
    geom_point(aes(y = post_mean),  color = "#CC5500",
             data = . %>% filter(State == "")) +
  geom_segment(aes(y = pq2.5, yend = pq97.5, xend = State), lineend='round',
               size = .4, color = "#DAA06D", data = . %>% filter(State == "")) +
  geom_segment(aes(y = pq25, yend = pq75, xend = State), lineend='round',
                color = "#CC5500", data = . %>% filter(State == "")) +
  ylab("Average reduction in gun homicides per 100,000") +
  xlab("Held out State") +
  facet_grid(loo ~ ., space = "free", scales = "free_y") +
  coord_flip() +
  theme_bw() +
  theme(strip.background = element_blank(),
        strip.text = element_blank(),
        axis.ticks.y = element_blank())

```


## Fig C.14: Multi-output correlation

```{r read_data_transform_mo}

start_year <- 1997
end_year <- 2018
t_int <- 2007
gunrate_states <- read_dta("ucrpanel2.dta") %>% 
  filter(year >= start_year, year <= end_year, State != "DC") %>% 
  group_by(State) %>% 
  summarise(hasguns=sum(nongunrate)) %>% 
  drop_na() %>% 
  pull(State)
# put the data into long format
read_dta("ucrpanel2.dta") %>% 
  filter(State != "DC") %>% 
  filter(year >= start_year) %>% 
  filter(year <= end_year) %>% 
  select(year, State, Population, gunrate, nongunrate) %>%
  pivot_longer(cols=c('gunrate', 'nongunrate'), names_to="crime_type") %>%
  filter(State %in% gunrate_states) %>% 
  drop_na() %>%
  mutate(
    treated=(State == "CA") * (year >= 2007),
    inv_pop=1/(Population/1e5),
    value_per=(value)
  ) -> long_df
num_years = end_year - start_year + 1
num_units = length(unique(long_df$State))
num_outcomes = 2
year_vector = start_year:end_year
# shape into a three dimensional array that is [outcome, year, unit]
murder_assault_arr = aperm(
  array(as.matrix(long_df  %>% select(value_per)), dim = c(num_years, num_units, 2)), 
  c(3, 1, 2)
)
control_idx = which(
  as.matrix(long_df %>% filter(crime_type=='gunrate') %>% select(treated)) == 0
)
inv_population = matrix(as.matrix((
  long_df %>% filter(crime_type == 'gunrate') %>% select(inv_pop) 
)), num_years, num_units)
```


### Functions
```{r}

organize_post_samples <- function(samples, unit, time, data, noutcomes=1) {
  time <- enquo(time)
  unit <- enquo(unit)
  times <- data %>% distinct(!!time) %>% pull(!!time)
  units <- data %>% distinct(!!unit) %>% pull(!!unit)
  
  # get time-unit-sample pairs
  time <- as_label(time)
  unit <- as_label(unit)
  times_units_samples <- expand.grid(1:max(samples$.draw), units, times, 1:noutcomes)
  names(times_units_samples) <- c("sample", unit, time, "crime_type")
  times_units_samples <- times_units_samples %>% mutate(crime_type = ifelse(crime_type==1, "gunrate", "nongunrate"))
  times_units_samples %>% mutate(year=as.integer(year))
  data <- data %>% mutate(year == as.integer(year))
  times_units_samples <- cbind(times_units_samples, samples)

  return(left_join(times_units_samples,
                   data[, c("year", "State", "crime_type", "value", "inv_pop")],
                   by=c("year", "State", "crime_type")))
  
}

```


```{r gaus_summary_mo}
## Homoskedastic Gaussian
# res_gaus <- readRDS("../../results/mtgp_fits_homoskedastic.RDS")

res <- readRDS("../results/mo_sweep.rds")

all_samples <- map_dfr(names(res), function(rank_name) {
  print(rank_name)
  fit <- res[[rank_name]]
  draws <- as_draws_matrix(fit)
  samps <- draws %>% spread_draws(f[M, N, D], f_samples[M, N, D], sigman[M])
  all_samples <- organize_post_samples(samps, unit = State,
                        time = year,
                        data = long_df %>% filter(year >= start_year),
                        noutcomes = max(samps$M))

  all_samples$rank <- rank_name
  all_samples
})

compute_summary_gaus <- function(all_samples, summary_function, group_vars = vars(sample, crime_type, rank), filters = function(...){ filter(..., year < t_int, State == "CA") }) {
  
  results_table <- all_samples %>%
    filters %>% 
    group_by(!!!group_vars) %>%
    summary_function

    results_table
}

```

```{r mo_cor}

cor_fn <- function(...){ 
  pivot_wider(..., id_cols = c("sample", "State", "year", "rank"), names_from="crime_type", names_sep = ".", values_from = c(f, f_samples, value)) %>%
  summarize(cor_obs = cor(value.gunrate, value.nongunrate), 
            cor_ppd = cor(f_samples.gunrate, f_samples.nongunrate)) }

cor_table_ca <- compute_summary_gaus(all_samples %>% filter(rank == 5), cor_fn, group_vars=vars(sample, State, rank))


cor_table_ca %>% ggplot() + geom_density(aes(x=cor_ppd), fill="gray") + xlim(c(-1, 1)) + theme_bw(base_size=16) + ggtitle("Posterior Predictive Correlation", subtitle="between Gun and Non-gun related homicides (California)") + xlab("Correlation") + ylab("")
#+ geom_vline(aes(xintercept=cor_obs), linetype="dashed", col="blue")

```

